---
title: "simpleRegress"
author: "Valerie Rodriguez"
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

This graph tells us that the longer the bear gets the wider the neck gets. The residual standard error is 5.4 and the adjusted R-squared is 0.7476.

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

This graph explains that as the length of the bear grows it gets older. The residual standard error is 7.5 and the adjusted R-squared is 0.5073.


```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

This graph shows that the longer the bear grows so does the length of the head.  The residual standard error is 4.2 and the adjusted R-squared is 0.8434. So far this one has the largest r squared.

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

This graph shows that head width grows as length of bear grows. The residual standard error is 7.3 and the adjusted R-squared is 0.5321.


```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

This graph shows that the chest of the bear grows as the length grows. The residual standard error is 4.9 and the adjusted R-squared is 0.7869.

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

This graph shows that the weight of the bear grows as the length grows. The residual standard error is 5.43 and the adjusted R-squared is 0.7423.


##compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsqared(bigger is better), or ny RSE (residual standard error) where smaller is better.

I will be ordering according to adjusted-Rquared from best to worst.


## now predict the length of a bear who has a neck of 17, 22 and 26


## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

